---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

## Introduction
The MAUT models are defined with aid of utility functions $u_1,\ldots,u_n$ which are evaluated over
indexes $x_1,\ldots,x_n$ and those utilities are aggregated considering additional weights 
$w_1,\ldots,w_n$, the whole final utility is given by the addition

\[u(x_1,\ldots,x_n) = \sum_{1\leq i \leq n}\, w_i\, u_i\ ( x_i )\]

With **mau** you can build and test decision models based in Multi Attribute Utility
Theory (MAUT). The utilities of any level of the decision model can be easily evaluated.

## Installation
To instal **mau** you can proceed in the following way making use of the devtools library
```{r, eval=FALSE}
library( devtools )
install_github( "pedroguarderas/mau" )
```

## Standard utility definition
The utility functions for a MAUT model could be defined in a practical format when the utilities
are piecewise defined like constant risk adverse functions, in such case it is only necessary 
to define the parameters of the function for each part of the domain of definition. This is because
the constant risk adversion functions are only of the following form $u(x) = a \cdot x + b$ or 
$u(x) = a \cdot e^{b \cdot x} + c$. Then the format for the piecewise defintion of the utility is    

>Header  
>  
>Function name  
>min_1 max_1 a_1 b_1 c_1  
>min_2 max_2 a_2 b_2 c_2  
>min_3 max_3 a_3 b_3 c_3  
>...  
>Function name  
>min_1 max_1 a_1 b_1 c_1  
>min_2 max_2 a_2 b_2 c_2  
>min_3 max_3 a_3 b_3 c_3  
>...  

If $c_i$ is $0$ then the utility is linear, otherwise is an exponential function. For example:
```{r, eval=TRUE}
library( mau )
file<-system.file("extdata", "utilities.txt", package = "mau" )
lines<-readLines( file )
for ( i in 1:length( lines ) ) { 
  cat( lines[i], '\n' )
}
```

## Main example
In the sources below is developed a complete example of a MAUT model, the package **mau** is 
employed to load utilities defined in the file `utilities.txt`, automatically an script with 
utilies is built and saved in the local working directory, after that with `Eval.Utilities` every 
function is evaluated over the columns of the index table, the names for utilities were previously
standarized with `Stand.String`. With another file `tree.csv` the decision tree associated to the
MAUT model is built and every weight and relative weight assigned with the `Make.Decision.Tree`
function, in addition the whole model with utilies of every criteria is obtained with 
`Compute.Model`. The simulation of constrained weights is made with `Sim.Const.Weights`, the result
could be employed for a sensitivy test of the decision model regarding concentrated weights 
variation.
```{r,eval=FALSE}
# Loading packages --------------------------------------------------------------------------------
library( mau )
library( data.table )
library( igraph )
library( ggplot2 )

# Table of indexes --------------------------------------------------------------------------------
index<-data.table( cod = paste( 'A', 1:10, sep = '' ), 
                   i1 = c( 0.34, 1, 1, 1, 1, 0.2, 0.7, 0.5, 0.11, 0.8 ),
                   i2 = c( 0.5, 0.5, 1, 0.5, 0.3, 0.1, 0.4, 0.13, 1, 0.74 ), 
                   i3 = c( 0.5, 1.0, 0.75, 0.25, 0.1, 0.38, 0.57, 0.97, 0.3, 0.76 ),
                   i4 = c( 0, 0.26, 0.67, 0.74, 0.84, 0.85, 0.74, 0.65, 0.37, 0.92 ) )

# Loading utilities -------------------------------------------------------------------------------
file<-system.file("extdata", "utilities.txt", package = "mau" )
script<-'utilities.R'
lines<-17
skip<-2
encoding<-'utf-8'
functions<-Read.Utilities( file, script, lines, skip, encoding )
source( 'utilities.R' )

# Index positions ---------------------------------------------------------------------------------
columns<-c( 2, 3, 4, 5 )

# Function names
functions<-sapply( c( 'Project', 
                      'Self implementation',
                      'External and local relations', 
                      'Scope of capabilities' ),
                   FUN = Stand.String )
names( functions )<-NULL

# Evaluation of utilities -------------------------------------------------------------------------
utilities<-Eval.Utilities( index, columns, functions )

# Tree creation -----------------------------------------------------------------------------------
file<-system.file("extdata", "tree.csv", package = "mau" )
sheetIndex<-1
cols<-1:5
rows<-c(1,8)
tree.data<-Read.Tree( file, skip = 0, nrow = 8 )
tree<-Make.Decision.Tree( tree.data )

# Compute the decision model ----------------------------------------------------------------------
weights<-tree.data[ !is.na( weight ) ]$weight
model<-Compute.Model( tree, utilities, weights )

# Weights simulation ------------------------------------------------------------------------------
n<-200
alpha<-c( 0.2, 0.5, 0.1, 0.2 )
constraints<-list( list( c(1,2), 0.7 ), 
                   list( c(3,4), 0.3 ) )
S<-Sim.Const.Weights( n, utilities, alpha, constraints )
plot.S<-Plot.Simulation.Weight( S$simulation, title = 'Simulations', 
                                xlab = 'ID', ylab = 'Utility' ) 
plot( plot.S )
```
