---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```
## Introduction
With **mau** you can build and test decision models based in Multi Attribute Utility
Theory (MAUT). With this package you can automatically evaluate utilities over a data.table of
indexes. The criteria of the decision tree could evaluated and easily computed.

The MAUT models are defined with aid of utility functions $u_1,\ldots,u_n$ which are evaluated over
indexes $x_1,\ldots,x_n$ and those utilities are aggregated considering additional weights 
$w_1,\ldots,w_n$, the whole final utility is given by the addition 
$u(x_1,\ldots,x_n) = \sum_{i=1,\ldots,n}\, w_i\, u_i( x_i )$.

With **mau** the decision model can be computed easily for any type of single valueted utility 
function. In addition, the utility functions could be defined considering the constant risk 
aversion criteria sujested by Arrow, those functions are only of two 
kinds $u(x) = a \cdot x + b$ or $u(x) = a \cdot e^{b \cdot x} + c$.

## Installation
To instal **mau** you can proceed in the following way making use of the devtools library
```{r, eval=FALSE}
library( devtools )
install_github( "pedroguarderas/mau" )
```

## Main example
In the sources below is developed a complete example of a MAUT model, the package **mau** is 
employed to load utilities defined in the file `utilities.txt`, automatically an script with 
utilies is built and saved in the local working directory, after that with `Eval.Utilities` every 
function is evaluated over the columns of the index table, the names for utilities were previously
standarized with `Stand.String`. With another file `tree.csv` the decision tree associated to the
MAUT model is built and every weight and relative weight assigned with the `Make.Decision.Tree`
function, in addition the whole model with utilies of every criteria is obtained with 
`Compute.Model`. The simulation of constrained weights is made with `Sim.Const.Weights`, the result
could be employed for a sensitivy test of the decision model regarding concentrated weights 
variation.
```{r,eval=FALSE}
# Loading packages --------------------------------------------------------------------------------
library( mau )
library( data.table )
library( igraph )
library( ggplot2 )

# Table of indexes --------------------------------------------------------------------------------
index<-data.table( cod = paste( 'A', 1:10, sep = '' ), 
                   i1 = c( 0.34, 1, 1, 1, 1, 0.2, 0.7, 0.5, 0.11, 0.8 ),
                   i2 = c( 0.5, 0.5, 1, 0.5, 0.3, 0.1, 0.4, 0.13, 1, 0.74 ), 
                   i3 = c( 0.5, 1.0, 0.75, 0.25, 0.1, 0.38, 0.57, 0.97, 0.3, 0.76 ),
                   i4 = c( 0, 0.26, 0.67, 0.74, 0.84, 0.85, 0.74, 0.65, 0.37, 0.92 ) )

# Loading utilities -------------------------------------------------------------------------------
file<-system.file("extdata", "utilities.txt", package = "mau" )
script<-'utilities.R'
lines<-17
skip<-2
encoding<-'utf-8'
functions<-Read.Utilities( file, script, lines, skip, encoding )
source( 'utilities.R' )

# Index positions ---------------------------------------------------------------------------------
columns<-c( 2, 3, 4, 5 )

# Function names
functions<-sapply( c( 'Project', 
                      'Self implementation',
                      'External and local relations', 
                      'Scope of capabilities' ),
                   FUN = Stand.String )
names( functions )<-NULL

# Evaluation of utilities -------------------------------------------------------------------------
utilities<-Eval.Utilities( index, columns, functions )

# Tree creation -----------------------------------------------------------------------------------
file<-system.file("extdata", "tree.csv", package = "mau" )
sheetIndex<-1
cols<-1:5
rows<-c(1,8)
tree.data<-Read.Tree( file, skip = 0, nrow = 8 )
tree<-Make.Decision.Tree( tree.data )

# Compute the decision model ----------------------------------------------------------------------
weights<-tree.data[ !is.na( weight ) ]$weight
model<-Compute.Model( tree, utilities, weights )

# Weights simulation ------------------------------------------------------------------------------
n<-200
alpha<-c( 0.2, 0.5, 0.1, 0.2 )
constraints<-list( list( c(1,2), 0.7 ), 
                   list( c(3,4), 0.3 ) )
S<-Sim.Const.Weights( n, utilities, alpha, constraints )
plot.S<-Plot.Simulation.Weight( S$simulation, title = 'Simulations', 
                                xlab = 'ID', ylab = 'Utility' ) 
plot( plot.S )
```
