---
title: "Rankings"
author: "Pedro Guarderas"
date: "October 25, 2023"
output: html_vignette
vignette: >
  %\VignetteIndexEntry{Rankings}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Colley method
```{r,eval=TRUE,message=FALSE, warning=FALSE}
library( mau )
```

```{r, eval=TRUE}
# Example without ties
d <- 10
n <- matrix( sample( x = 0:5, size = d * d, replace = TRUE ), d, d )
n <- n + t( n )
diag( n ) <- 0
g <- rowSums( n )
# Number of win matches for each team
w <- sapply( 1:d, FUN = function( i ) sample( x = 1:g[i], size = 1, replace = TRUE ) )
# Number of lost matches for ech team
l <- rowSums( n ) - w
r <- colley_rating( n, w, l )
print( r )
```

## Offensive - Defensive rating method
```{r, eval=TRUE}
A <- matrix( c( 0, 7, 21, 7, 0,
                52, 0, 34, 25, 27, 
                24, 16, 0, 7, 3,
                38, 17, 5, 0, 14,
                45, 7, 30, 52, 0 ), 
             nrow = 5, ncol = 5 )
r <- od_rating( A )
print( r )
```

## Borda counts
We take $C$ the finite set of options of candidates, ranking matrix $R \in C^{m\times n}$, number of 
votes for each case $v \in \mathbb{R}^n$, $u \in \mathbb{R}^m$ weights for each rank, $N = |C|$ 
number of options or candidates, $l : C \longrightarrow \{1, \ldots, N\}$ labels by option, 
$W \in \mathbb{R}^{N \times n}$ the ranking matrix. For each $i \in \{1, \ldots, N\}$ and 
$j \in \{1, \ldots, n\}$ we compute
$$
k = l( R_{i,j} ), \qquad W_{k,j} = u_i
$$

$\text{order}$ returns the order index of a vector.
$$
w = W v,\qquad r = \text{order}( w )
$$


```{r, eval=TRUE}
# Position of each participant
R <- matrix( c( 2, 3, 1, 3, 1, 2, 1, 2, 3 ), nrow = 3, ncol = 3 )
v <- c( 7, 5, 2 )
bc <- borda_count( R, v )
print( bc )
```

```{r, eval=TRUE}
R <- matrix( c( 'A', 'B', 'C', 'B', 'A', 'C', 'C', 'A', 'B' ), nrow = 3, ncol = 3 )
v <- c( 5, 4, 3 )
bc <- borda_count( R, v )
print( bc )
```

```{r, eval=TRUE}
R <- matrix( c( 'Steak', 'Chicken', 'Barbacoa', 'Carnitas', 
                'Carnitas', 'Barbacoa', 'Chicken', 'Steak', 
                'Chicken', 'Carnitas', 'Steak', 'Barbacoa',
                'Steak', 'Barbacoa', 'Chicken', 'Carnitas' ), nrow = 4, ncol = 4 )
v <- c( 8, 7, 5, 3 )
bc <- borda_count( R, v )
print( bc )
```

```{r, eval=TRUE}
R <- matrix( c( 'A', 'B', 'C', 'B', 'A', 'C', 'C', 'A', 'B' ), nrow = 3, ncol = 3 )
v <- c( 5, 4, 3 )
u <- c( 6, 4, 2 )
bc <- borda_count( R, v, u )
print( bc )
```